---
title: "tidycensus attempt"
author: "JP"
date: "5/12/2019"
output: html_document
---

```{r message=FALSE}
library(magrittr)
library(scales)
library(schoolfundr)
library(tidycensus)
library(tidyverse)

source_file_paths()
source(here::here("strings.R"))
```

```{r}
ds_vars_dec_2010_cd113 <- load_variables(2010, "dec/cd113", cache = F) 

# View(ds_vars_dec_2010_cd113)

ds_vars_rent <- 
  ds_vars_dec_2010_cd113 %>% 
  filter(
    str_detect( 
      label, 
      regex("rent", ignore_case = TRUE)
    )
  )
```

Wondering what !! and Tenure refer to. "Total!!Renter occupied" and concept == "TENURE", which I assume means the number "units" occupied by renters because the definition of tenure focusses on *units* rather than *people* (see below). 

From [Census dictionary](https://www.census.gov/housing/hvs/definitions.pdf)
"**Tenure.** A unit is owner occupied if the owner or co-owner lives in the unit, even if it is mortgaged or not fully paid for. A cooperative or condominium unit is "owner occupied" only if the owner or co-owner lives in it. All other occupied units are classified as "renter occupied," including units rented for cash rent and those occupied without payment of cash rent."

## rental population by census tract in CA
```{r}
ds_vars_dec_2010_cd113 %>% 
  filter(str_detect(concept, "TOTAL POPULATION")) %>% 
  filter(str_detect(label, "Owned")) %>% count(label)

# H010001 # TOTAL POPULATION IN OCCUPIED HOUSING UNITS
# P001001 # Total population (may include homeless?)
# How are the institutionalized claimed? 

my_vars <- 
  c(total = "H010001", renters = "H011004", own_mortgage = "H011002", own_outright = "H011003")
```

```{r}
ds_ca <- 
  get_decennial(
    geography = "state", state = "CA", 
    variables = my_vars, year = 2010
  ) %>% 
  mutate(
    var_agg = case_when(
      str_detect(variable, "^own") ~ "owner",
      TRUE ~ variable
    )
  ) %>% 
  group_by(var_agg) %>% 
  summarise(value = sum(value)) 

# Check
ds_ca %>% 
  arrange(-value) %>% 
  mutate(check = first(value) - value)
```

## rental population by census tract in CA
```{r}
ds_ca_tract <- 
  get_decennial(
    geography = "tract", state = "CA", 
    variables = my_vars, year = 2010,
    geometry = TRUE
  ) 

ds_ca_tract %<>% 
  mutate(
    variable = case_when(
      str_detect(variable, "^own") ~ "owner",
      TRUE ~ variable
    )
  ) 

ds_ca_tract %<>% 
  group_by(NAME, variable) %>% 
    mutate(value = sum(value)) %>% 
  ungroup() %>% 
  unique()

ds_ca_tract %<>% 
  separate(NAME, into = c("tract", "county", "state"), sep = ",", extra = "warn") %>% 
  mutate(
    county = str_to_lower(county) %>% str_remove_all("county") %>% str_trim(),
    census_tract = str_remove_all(tract, "Census Tract") 
  ) %>% 
  filter(county %in% bay_area_counties) %>%  
  select(GEOID, census_tract, county, pop_group = variable, population = value, geometry)
```

```{r}
ds_ca_tract %<>% 
  filter(pop_group != "total") %>% 
  group_by(census_tract) %>% 
    mutate(percent = population / sum(population)) %>% 
  ungroup()
```

```{r}
ds_ca_tract %>% 
  filter(
    pop_group == "renters", 
    county %in% c("san mateo", "santa clara")
  ) %>% 
  ggplot(aes(fill = percent)) + 
  geom_sf(color = NA) + 
  scale_fill_viridis_c(
    option = "magma", 
    breaks = pretty_breaks(5), 
    labels = percent_format()
  )
```

```{r out.width="100%}
ds_ca_tract %>% 
  filter(
    pop_group == "owner", 
    county == "san mateo",
    !str_detect(census_tract, "6135.01"),
    !str_detect(census_tract, "6135.02"),
    !str_detect(census_tract, "6136"),
    !str_detect(census_tract, "6031"),
    !str_detect(census_tract, "6032"),
    !str_detect(census_tract, "6033"),
    !str_detect(census_tract, "6034"),
    !str_detect(census_tract, "6132"),
    !str_detect(census_tract, "6134"),
    !str_detect(census_tract, "6137"),
    !str_detect(census_tract, "6138")
  ) %>% 
  ggplot(aes(fill = percent)) + 
  geom_sf(color = NA) + 
  scale_fill_viridis_c(
    option = "magma", 
    breaks = pretty_breaks(5), 
    labels = percent_format()
  ) + 
  labs(fill = "Percent\nOwners") # + 
  # geom_sf_text(aes(label = census_tract), color = "gray")
```

# Can I add town outlines to the above map? 
Explore https://www.r-spatial.org/r/2018/10/25/ggplot2-sf-2.html.
```{r}

```

For help with tidycensus package see: 
https://github.com/walkerke/tidycensus/issues/129
https://mattherman.info/blog/tidycensus-mult/
https://notary.cdn.sos.ca.gov/forms/notary-county-codes.pdf